Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-07 alloscore_manuscript/supplement.rnw
give some refs for quantiles as decisions

We recall how quantiles arise as solutions to a probabilistic decision problem. Let $Y$ be a random variable
representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}_+$ be a decision variable
representing levels of some costly counter-measure, such as procurement of monoclonal antibody treatments,
that can be taken at a cost $C>0$ per unit in preparation for $Y$.\footnote{
Quantiles could also be derived for a problem in which $x$ and $Y$ take negative values, corresponding, for instance, to a decision
maker that operates in a resource market and a that $Y$ takes negative values when ``recoveries'' outnumber incidence.  But we do not consider
such scenarios in this work.	 
}  A decision maker must decide on a level $x$ of investment in the
counter-measure, and wishes to avoid excesses in either the expediture $Cx$ or the shortage $(y-x)_{+}$ when $Y=y$ is
realized. To formalize the trade-off between these potential excesses we quantify the loss associated with a unit of
shortage by a constant $L>C$ (which assumes that the counter-measure has some practical value) and combine the total shortage loss with expenditure into a \emph{loss function}\footnote
{This does involve a confusing use of the word \emph{loss} to refer to two different quantities, but this seems to be
an ingrained and unavoidable habit in the literature.}
\[
l(x,y) = Cx + L(y-x)_{+}.
\] 
The decision problem is then to select a random future loss $l(x,Y)$ in a way that aligns with the preference that $l(x,y)$ be
as low as possible given any realization $Y=y$.

>>>>

We recall how quantiles arise as solutions to a probabilistic decision problem, drawing on perspectives developed in the fields of
both forecasting (see e.g., \cite{gneiting2011quantiles} and \cite{jose2009evaluating}) and stochastic optimization 
(se e.g., \cite{royset2022optimization}, sections 1.C and 3.C).

Let $Y$ be a random variable
representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}_+$ be a decision variable
representing levels of some costly counter-measure, such as procurement of monoclonal antibody treatments,
that can be taken at a cost $C>0$ per unit in preparation for $Y$.\footnote{
Quantiles could also be derived for a problem in which $x$ and $Y$ take negative values, corresponding, for instance, to a decision
maker that both buys and sells in a resource market and a $Y$ that takes negative values when ``recoveries'' outnumber incidence.  
But we do not consider such scenarios in this work.	 
}  A decision maker must decide on a level $x$ of investment in the
counter-measure, and wishes to avoid excesses in either the expediture $Cx$ or the shortage $(y-x)_{+}$ when $Y=y$ is
realized. To formalize the trade-off between these potential excesses we quantify the loss associated with a unit of
shortage by a constant $L>C$ (which assumes that the counter-measure has some practical value) and combine the total shortage loss with expenditure into a \emph{loss function}\footnote
{This does involve a confusing use of the word \emph{loss} to refer to two different quantities, but this seems to be
an ingrained and unavoidable habit in the literature.}
\[
l(x,y) = Cx + L(y-x)_{+}.
\] 
The decision problem is then to select a random future loss $l(x,Y)$ in a way that aligns with the preference that $l(x,y)$ be
as low as possible given any realization $Y=y$.
